---
title: "Bioirrigation fitting examples"
author: "Emil De Borger"
date: "11-6-2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load uranine model.
install.packages("irrfit.tar", repos = NULL, type = "tar")
require(irrfit)
require(FME)
source("./bioirrigation_fitting_functions.R")
```

# Looking at tracer timeseries

Three examples of tracer data from measurements performed in the Dortsman intertidal area (Netherlands).

### First, the data

Uranine concentrations have been converted from FTU's (raw output of the sensor) to microgram/L using a concentration series performed 
before the experiment to calibrate the sensors.
The time is converted to hours, as that is how the model is set up.

```{r, echo = FALSE}

dortsman_data           <- read.table("./example_data/example_data_dortsman.csv", sep = ";", header = TRUE, dec = ".")
colnames(dortsman_data) <- c("time", "August", "November", "July")

cols <- gray.colors(3)
plot(dortsman_data$time , dortsman_data$August   , lwd = 3, col = cols[1], type = "l", xlab = "Time (h)", ylab = "Concentration (microgram/L)", ylim = c(3, 11))
lines(dortsman_data$time, dortsman_data$November , lwd = 3, col = cols[2])
lines(dortsman_data$time, dortsman_data$July, lwd = 3, col = cols[3])
legend("topright", c("August", "November", "July"), bty = "n", lty = 1, col = rev(gray.colors(3)), lwd = 3)

```

### Some corrections

There are some spikes in the beginning of the data, and it's best to cut this off to improve the fitting procedure.
The August data is fine, but in the September - November data some of the first observations are removed.
First, the data are subdivided in individual measurements, with specific column names, so it can be recognized by the fitting procedure. "UrOw" stands for uranine in overlying water.

```{r}

August   <- data.frame("time" = dortsman_data$time, "UrOW" = dortsman_data$August   )
November <- data.frame("time" = dortsman_data$time, "UrOW" = dortsman_data$November )
July     <- data.frame("time" = dortsman_data$time, "UrOW" = dortsman_data$July)

November <- November[-c(1:4),]
November <- November[!(November$time >= 21.4),]
July     <- July[-(1:4),]

```

Now check again:

```{r}

plot(August$time    , August$UrOW   , lwd = 3, col = 2, type = "l", xlab = "Time (h)", ylab = "Concentration (microgram/L)", ylim = c(3, 11))
lines(November$time , November$UrOW , lwd = 3, col = 3)
lines(July$time, July$UrOW, lwd = 3, col = 4)
legend("topright", c("August", "November", "July"), bty = "n", lty = 1, col = 2:4, lwd = 3)

```

# Preparing parameters

```{r}

# There are three sets of parameters that are specified (in this case).
# Base parameters that are the same in all months.
sediment_parms <- c(kAds = 1, dens = 2.4952, porosity = 0.429, eqAds = .05, pumpSurfDepth = 0.001)

# Parameters specific to each core
August_specific   <- c(hOw = 12  , hSed = 18  , iniConcOW = 5.9293, area = 314.15)
November_specific <- c(hOw = 13.5, hSed = 16.5, iniConcOW = 7.6626, area = 314.15)
July_specific     <- c(hOw = 8.5 , hSed = 21.5, iniConcOW = 10.70263, area = 314.15)

# Combining both
parameters_August    <- c(August_specific  , sediment_parms) 
parameters_November  <- c(November_specific, sediment_parms) 
parameters_July      <- c(July_specific    , sediment_parms)

# Initial starting parameters for parameters to fit
fitp <- c(pumpSurfRate = 2, pumpSurfAtt = 1)

```

# First run of the model

Without fitting. Now we see why further tuning of the irrigation parameters is necessary.

```{r}

out_August   <- run.Uranine(parms = c(parameters_August  , fitp), hours = 24)
out_November <- run.Uranine(parms = c(parameters_November, fitp), hours = 24)
out_July     <- run.Uranine(parms = c(parameters_July    , fitp), hours = 24)

par(mfrow = c(1,1))
plot(out_August, out_November, out_July, obs = list(August, November, July), col = 2:4, ylim = c(3, 11),
     obspar = list(cex = 4, col = 2:4, pch = "."))

```

# Fitting

* Please check the vignette of the package FME `citation("FME")` to learn more about the fitting procedure used here..

```{r}

Fit_August    <- modFit(f = URAcost, p = fitp, corepars = parameters_August, 
                    measured.data = August, verbose = FALSE,
                    lower=c(0.001, 0.001),
                    upper=c(20, 10),
                    method = "Pseudo",
                    control = list(numiter = 5000))
Fit_November  <- modFit(f = URAcost, p = fitp, corepars = parameters_November,
                    measured.data = November, verbose = FALSE,
                    lower=c(0.001, 0.001),
                    upper=c(5, 10),
                    method = "Pseudo",
                    control = list(numiter = 10000))
Fit_July     <- modFit(f = URAcost, p = fitp, corepars = parameters_July,
                    measured.data = July, verbose = FALSE,
                    lower=c(0.001, 0.001),
                    upper=c(20, 10),
                    method = "Pseudo",
                    control = list(numiter = 10000))

```

# Plot the fitted output

```{r}

## First run the model with the fitted parameters as input.
final_August   <- run.Uranine(parms = c(coef(Fit_August)  , parameters_August)   , hours = 24, times = seq(from = 0, to = 24, length.out = 288), rtol = 1e-8)
final_November <- run.Uranine(parms = c(coef(Fit_November), parameters_November) , hours = 24, times = seq(from = 0, to = 24, length.out = 288), rtol = 1e-8)
final_July     <- run.Uranine(parms = c(coef(Fit_July)    , parameters_July), hours = 24, times = seq(from = 0, to = 24, length.out = 288), rtol = 1e-8)

plot(August, xlab ="time (hours)", ylab = "UrOW (ug/L)", ylim = c(2,8), xlim = c(0,20), main = "August", cex = 0.6, col = "grey")
lines(out_August[,1], out_August[,2], lty=2)
lines(final_August[,1], final_August[,2], col = 2, lwd = 1)
legend("topright", legend = c("Measured data", "Initial run", "Fitted run"), lty = c(1, 2, 1), col = c("grey",1,2), lwd = c(3,1,1), bty = "n")

plot(November, xlab ="time (hours)", ylab = "UrOW (ug/L)", ylim = c(0,10), xlim = c(0,25), main = "November", col = "grey")
lines(out_November[,1], out_November[,2], lty=2)
lines(final_November[,1], final_November[,2], col = 3, lwd = 1)
legend("topright", legend = c("Measured data", "Initial run", "Fitted run"), lty = c(1, 2, 1), col = c("grey",1,3), lwd = c(3,1,1), bty = "n")

plot(July, xlab ="time (hours)", ylab = "UrOW (ug/L)", ylim = c(6,11), xlim = c(0,20), main = "July", col = "grey")
lines(out_July[,1]  , out_July[,2], lty=2)
lines(final_July[,1], final_July[,2], col = 4, lwd = 1)
legend("topright", legend = c("Measured data", "Initial run", "Fitted run"), lty = c(1, 2, 1), col = c("grey",1,4), lwd = c(3,1,1), bty = "n")

```

# The coefficients

```{r}
# August
print(coef(Fit_August))

# November
print(coef(Fit_November))

# September
print(coef(Fit_July))

```